Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-14 chi2012.tex

While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other Òdata specialistsÓ remains minimal.  Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.  For example, while much of this data represents a many and varied set of things, activities, and descriptions at various granularities, most commonly used structured personal information management (PIM) tools either can only manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.

@@

%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

\subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.  We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).  This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant's experiences, while collecting the same general areas of information.  Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.

@@

Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.  They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.   They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.  People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.  They considered how to get there and looked up train or bus times on the Web.  One participant organised carpools informally and at last minute via Facebook or text messages.

When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.  Also, their friend's preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.  Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.

@@

\subsection{PS.2 - Technical Challenges of Integrating the Data}
Our investigation of integration challenges started with identifying a set of data sources to analyze.  For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.  In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.   Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.

We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.  Based on statistics collected We first selected 4 different categories

%% TODO THIS IS PLACEHOLDER DATA >> 
\begin{table*}
\small
\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l |}
\hline
source category  & data type & sources & API type & avg fields/rec & avg depth & term. het & struc. het \\
\hline
social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
event calendars & events & eventbrit, gcal  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
retail (shopping)  & product profile & Amazon, shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
music catalogues & songs and albums & echonest, rdio, soundcloud, spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
weather & forecasts & wunderground, weatherbug, yahoo weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
\hline
\end{tabular}
\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed} \label{tbl_tps}
\end{table*}
%% <<

@@

\subsection{Participants}
We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.  All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.  Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.  All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.

@@

\subsection{Data Collection}
Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.  During the lab study, we recorded the audio and the participant's actions on screen. We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.

The studies lasted 40 minutes, on average.  Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.  The studies were run over 4 days.  At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.

@@

During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.  After we had collected the study's data, we performed our data analysis.

@@

%What we found 
%
%What we didn't find (evidence for)
%
%What we might look at next -- follow-on work
%		We focused on terminological heterogeneity -- didn't address structural heterogeneity, semantic heterogeneity
%		We focused on finite (small) sets Streaming data (how do select future data --) 
%		Scale - small data, what are the facilites required for large data
%		Collaborative - how might interaction support multiple users collaborating?
%
%Limitations found from study... what is the take away from the limitations

>>>>

While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other “data specialists” remains minimal.  Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.  For example, while much of this data represents a many and varied set of things, activities, and descriptions at various granularities, most commonly used structured personal information management (PIM) tools either can only manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.

@@

In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.  For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).  We analysed the data collected from this study using a thematic approach.  From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).  We performed a statistical analysis of the features of the available data.  In the next two sections, we describe our two pre-studies.
%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

\subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
% \subsubsection{Method}
Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.  We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).  This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.  Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.

@@

Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.  They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.   They selected venues, locations, restaurants and activities based on their own knowledge, friend’s recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.  People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.  They considered how to get there and looked up train or bus times on the Web.  One participant organised carpools informally and at last minute via Facebook or text messages.

When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.  Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.  Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.

@@

\subsection{PS.2 - Technical Challenges of Integrating the Data}

The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.

\subsubsection{Method}
We sought out APIs and data feeds of popular Web services and sites that were aimed at developers targeting consumption of the data by end-users.  Based on statistics collected We first selected 4 different categories

@@

\subsection{Participants}
We recruited 20 participants (10 female) from the University of Southampton’s international student mailing list.  All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.  Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.  All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.

@@

\subsection{Data Collection}
Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study’s protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.  During the lab study, we recorded the audio and the participant’s actions on screen. We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.

The studies lasted 40 minutes, on average.  Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn’t influence the study.  The studies were run over 4 days.  At the end of each day the facilitators and observers met to discuss the studies’ data and to revisit protocols, if necessary.

@@

During the data collection we collated a spreadsheet summarising the participant’s gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.  After we had collected the study’s data, we performed our data analysis.

@@

%What we found 
%
%What we didn’t find (evidence for)
%
%What we might look at next -- follow-on work
%		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
%		We focused on finite (small) sets Streaming data (how do select future data --) 
%		Scale - small data, what are the facilites required for large data
%		Collaborative - how might interaction support multiple users collaborating?
%
%Limitations found from study... what is the take away from the limitations
